---
title: "Edit metadata templates"
output: html_document
---

```{r setup, include=FALSE}

# Parameterize

dir_path <- '/Users/csmith/Desktop/data_packages'
data_path <- '/Users/csmith/Desktop/data_packages/edi_301/data_objects'

# Initialize .Rmd

knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
library(EMLassemblyline)

# Clear contents of data package directory

unlink(
  x = '/Users/csmith/Desktop/data_packages/edi_301',
  recursive = TRUE
)

# Create directory structure

template_directories(
  path = dir_path,
  dir.name = 'edi_301'
)

# Add completed templates

file.copy(
  from = paste0(
    system.file(
      '/examples/templates',
      package = 'EMLassemblyline'
    ),
    '/',
    list.files(
      system.file(
        '/examples/templates',
        package = 'EMLassemblyline'
      )
    )
  ),
  to = paste0(
    dir_path,
    '/edi_301/metadata_templates'
  )
)

# Remove templates

file.remove(
  paste0(
    dir_path,
    '/edi_301/metadata_templates/bounding_boxes.txt'
  )
)

file.remove(
  paste0(
    dir_path,
    '/edi_301/metadata_templates/geographic_coverage.txt'
  )
)

file.remove(
  paste0(
    dir_path,
    '/edi_301/metadata_templates/taxonomicCoverage.xml'
  )
)

# Add data objects

file.copy(
  from = paste0(
    system.file(
      '/examples/data',
      package = 'EMLassemblyline'
    ),
    '/',
    list.files(
      system.file(
        '/examples/data',
        package = 'EMLassemblyline'
      )
    )
  ),
  to = paste0(
    dir_path,
    '/edi_301/data_objects'
  )
)

# Create templates

template_geographic_coverage(
  path = paste0(
    dir_path,
    '/edi_301/metadata_templates'
  ),
  data.path = paste0(
    dir_path,
    '/edi_301/data_objects'
  ),
  data.table = 'nitrogen.csv',
  site.col = 'site_name',
  lat.col = 'site_lat',
  lon.col = 'site_lon'
)

# Set working directory

knitr::opts_knit$set(root.dir = '/Users/csmith/Desktop/data_packages/edi_301')

```

## Editing

While a substantial portion of metadata content is auto-extracted, most templates require some user input. Template files can be edited in simple text and spreadsheet editors. Specific instructions for each template are found below.

_NOTE:_

*  _Some content in tabular templates is optional. Leave these spaces blank (no need to add NAs)._
*  _Only UTF-8 is supported by EML._

## abstract.txt

Like a journal article, the abstract provides a concise summary of the data package. It should cover what, why, when, where, and how.

Example:

```{r abstract.txt, comment='', echo = FALSE}

# Display template

abstract <- suppressWarnings(
  readLines('./metadata_templates/abstract.txt')
)

knitr::kable(
  row.names = FALSE, 
  col.names = '', 
  abstract, caption = 'abstract.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '200px')

```

## methods.txt

Describe the specific methods used to create the data. Include as much detail as necessary to allow a future scientist to assess fitness for use. Add instrument descriptions, protocols, etc. 

Example:

```{r methods.txt, comment='', echo = FALSE}

# Display template

methods <- suppressWarnings(
  readLines('./metadata_templates/methods.txt')
)

knitr::kable(
  row.names = FALSE, 
  col.names = '', 
  methods, caption = 'methods.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '200px')

```

## keywords.txt

Keywords facilitate search and discovery along scientific terms, as well as research groups, field stations, and other organizations. Using a controlled vocabulary or thesaurus vastly improves discovery (e.g. [LTER](http://vocab.lternet.edu/vocab/vocab/index.php).

__keywords.txt__ is a tab delimited table with columns:

*  **keyword** One keyword per row.
*  **keywordThesaurus** (optional) URI of controlled vocabulary from which the keyword was found.

Example:

```{r view keywords, echo=FALSE}

# Read keywords template

keywords <- read.table(
  file = './metadata_templates/keywords.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 2)
)

keywords <- keywords[keywords$keyword != '', ]

keywords[c(2,3,5), 'keywordThesaurus'] <- 'https://vocab.lternet.edu/vocab/vocab/index.php'

# Display template

knitr::kable(
  row.names = FALSE, 
  x = keywords, caption = 'keywords.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '60%', height = '300px')

```

## personnel.txt

Personnel and funding information facilitates attribution and credit for the data package.

__personnel.txt__ is a tab delimited table with columns:

* **givenName** First name
* **middleInitial** (optional) Middle initial
* **surName** Last name
* **organizationName** Organization the person is associated with
* **electronicMailAddress** Email address
* **userId** (optional) Person's unique identifier (e.g. [ORCID](https://orcid.org/)).
* **role** Person’s role with respect to the data package. Persons serving more than one role, or roles shared among more than one person, should be added as separate lines. Valid roles are:
    + **creator** Data package author
    + **PI** (optional) Principal investigator associated with the data package.
    + **contact** Data package contact. Can be an organization or position (e.g. data manager). To do this, enter the organization or position name under *givenName* and leave *middleInitial* and *surName* empty.
    + Any other roles are acceptable and will be listed under _associated party_ (e.g. Field Technician).
* **projectTitle** (optional) Project title the data package was created under. Only list project titles on lines where the personnel role is *PI*. If an ancillary project was involved, then add a new row below the primary project with the ancillary project title and associated PI. Repeat for additional ancillary projects.
* **fundingAgency** (optional) Agency the project was funded by. Only list on lines where the role is *PI*.
* **fundingNumber** (optional) Grant or award number funding the project. Only list on lines where the role is *PI*.

Example:

```{r view personnel, echo=FALSE}

# Read personnel template

personnel <- read.table(
  file = './metadata_templates/personnel.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 10)
)

personnel <- personnel[personnel$givenName != 'Joe', ]
personnel <- personnel[personnel$givenName != 'Chase', ]
personnel <- personnel[personnel$givenName != 'Organization lead', ]
personnel <- personnel[personnel$organizationName != 'Center for Limnology', ]
personnel <- personnel[personnel$organizationName != 'Trout Lake Station', ]

# Display template

knitr::kable(
  row.names = FALSE, 
  x = personnel, caption = 'personnel.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '400px')

```

## intellectual_rights.txt

The intellectual rights license describes how the data package may be used in the future. We strongly recommend releasing without restriction ([CC0](https://creativecommons.org/publicdomain/zero/1.0/)) or with minimal attribution ([CC BY](https://creativecommons.org/licenses/by/4.0/)).

Example:

```{r intellectual_rights.txt, comment='', echo = FALSE}

# Display template

intellectual_rights <- suppressWarnings(
  readLines('./metadata_templates/intellectual_rights.txt')
)

knitr::kable(
  row.names = FALSE, 
  col.names = '', 
  intellectual_rights, caption = 'intellectual_rights.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '200px')

```

## additional_info.txt

(*Optional*) Ancillary information about a data package not within the scope of abstract or methods (e.g. list of research articles created from this dataset).

Example:

```{r additional_info.txt, comment='', echo = FALSE}

# Display template

additional_info <- suppressWarnings(
  readLines('./metadata_templates/additional_info.txt')
)

knitr::kable(
  row.names = FALSE, 
  col.names = '', 
  additional_info, caption = 'additional_info.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '50%', height = '100px')


```

## attributes_*.txt

(*Required if data package contains data tables*) Well documented data table attributes facilitate human and machine understanding.

__attributes.txt__ is a tab delimited table with columns:

* **attributeName** Column name
* **attributeDefinition** Column definition
* **class** Column class. Valid options are:
    + **numeric** Numeric variable
    + **categorical** Categorical variable (i.e. nominal)
    + **character** Free text character variable (e.g. notes)
    + **Date** Date and time variable
* **unit** Column unit. Required for numeric classes. Select from the standard unit dictionary, accessible via the R console command `view_unit_dictionary()`. If not found, then define as a custom unit (see custom_units.txt).
* **dateTimeFormatString** Format string. Required for Date classes. Valid format string components are:
    + **Y** Year
    + **M** Month
    + **D** Day
    + **h** Hour
    + **m** Minute
    + **s** Second
All separators of format string components (e.g. - / \ :) are supported.
* **missingValueCode** Missing value code. Required for columns containing a missing value code).
* **missingValueCodeExplanation** Definition of missing value code.

Example:

```{r view attributes, echo=FALSE}

# Read attribute templates

attributes_decomp <- read.table(
  file = './metadata_templates/attributes_decomp.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 7)
)

attributes_nitrogen <- read.table(
  file = './metadata_templates/attributes_nitrogen.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 7)
)

# Display attribute templates

knitr::kable(
  attributes_decomp, caption = 'attributes_decomp.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '300px')


knitr::kable(
  attributes_nitrogen, caption = 'attributes_nitrogen.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '300px')

```

## custom_units.txt

(*Required if non-standard units are used in a data tables attribute.txt file*) Custom unit definition and relation to SI units facilitates human and machine understanding.

__custom_units.txt__ is a tab delimited table with columns:

* **id** Unit name (e.g. feetPerSecond)
* **unitType** Unit type (e.g. velocity)
* **parentSI** SI equivalent (e.g. metersPerSecond)
* **multiplierToSI** Multiplier to SI equivalent (e.g. 0.3048)
* **description** Abbreviation (e.g. ft/s)

Example:

```{r view custom_units, echo=FALSE}

# Read custom units template

custom_units <- read.table(
  file = './metadata_templates/custom_units.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 5)
)

# Display attribute templates

knitr::kable(
  custom_units, caption = 'custom_units.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '250px')

```

## catvars_*.txt

(*Required if an attributes table contains categorical class variables*) Categorical variables require definitions to enable full understanding of data table contents.

__catvars.txt__ is a tab delimited table with columns:

* **attributeName** Column name
* **attributeDefinition** Column definition
* **class** Column class. Valid options are

Example:

```{r view catvars, echo=FALSE}

# Read catvar templates

catvars_decomp <- read.table(
  file = './metadata_templates/catvars_decomp.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 3)
)

catvars_nitrogen <- read.table(
  file = './metadata_templates/catvars_nitrogen.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 3)
)

# Display catvar templates

knitr::kable(
  catvars_decomp, caption = 'catvars_decomp.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '45%', height = '300px')


knitr::kable(
  catvars_nitrogen, caption = 'catvars_nitrogen.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '45%', height = '300px')

```

## geographic_coverage.txt

(*Optional*) Documented sampling points and areas facilitate data package discovery based on geographic search criteria.

__geographic_coverage.txt__ is a tab delimited table with columns:

* **geographicDescription** Description of geographic point or area.
* **northBoundingCoordinate** North coordinate
* **southBoundingCoordinate** South coordinate
* **eastBoundingCoordinate** East coordinate
* **westBoundingCoordinate** West coordinate

Coordinates must be in decimal degrees and include a minus sign (-) for latitudes south of the equator and longitudes west of the prime meridian. For points, repeat latitude and longitude coordinates in respective north/south and east/west fields.

Example:

```{r view geographic_coverage.txt, echo=FALSE}

# Read geocoverage template

geocoverage <- read.table(
  file = './metadata_templates/geographic_coverage.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 5)
)

# Display geocoverage template

knitr::kable(
  geocoverage, caption = 'geographic_coverage.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '400px')

```

## taxonomic_coverage.txt

(*Optional*) Fully documented taxonomic ranks and authority IDs for each taxa enable data package discovery based on taxonomic search criteria.

__taxonomic_coverage.txt__ is a tab delimited table with columns:

* **taxa** Taxa name as it occurs in the data. Can be single word or species binomial.
* **scientific_name** (optional) Taxa’s scientific name found in an authority system (e.g. _Perca flavescens_)
* **scientific_authority_system** (optional) Authority system in which the taxa’s scientific name was found (e.g. ITIS).
* **scientific_authority_id** (optional) Taxa’s identifier in the authority system (e.g. 168469).
* **common_name** (optional) Taxa’s common name found in an authority system (e.g. Yellow Perch).
* **common_authority_system** (optional) Authority system in which the taxa’s common name was found (e.g. ITIS).
* **common_authority_id** (optional) Taxa’s identifier in the authority system (e.g. 168469).

Example:

```{r view taxonomic_coverage.txt, echo=FALSE}

# Read taxonomic coverage template

taxacoverage <- read.table(
  file = './metadata_templates/taxonomic_coverage.txt',
  header = TRUE,
  sep = '\t',
  quote = "\"",
  as.is = TRUE,
  comment.char = "",
  fill = T,
  colClasses = rep("character", 7)
)

# Display taxonomic coverage template

knitr::kable(
  taxacoverage, caption = 'taxonomic_coverage.txt') %>%
  kableExtra::kable_styling('bordered', font_size = 12) %>%
  kableExtra::scroll_box(width = '100%', height = '400px')

```
