---
title: "Introduction to emlAssemblyLine"
author: "Colin Smith"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to emlAssemblyLine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Introduction
What is EML and why is it good?
Problem/challenge.
How does the EML Assembly Line simplify this challenge?
A description of the assembly line (figure).
- templates
- extraction functions
- wrapper
The assembly line can be retooled to provide additional functionality and a custom EML product.

While at first this may seem daunting. Of course this is a new process. Really, you have already done the most challengin part. Collected and refined the data to a usable set. You have all the information, now log it through these steps. Give your self a couple days to work through all this.

## Current capabilities and requirements
Data tables, comma or tab delimited

Our priorities for building up the assembly line to produce EML for: spatial vectors and spatial raster
data.

## Collaboration
[Collaborate with us](https://github.com/EDIorg/emlAssemblyLine)


## References
[Ecological Metadata Language](https://knb.ecoinformatics.org/#external//emlparser/docs/index.html)

[EML R code package](https://github.com/ropensci/EML)

# Assembling EML for an example dataset.





Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
