---
title: "Organize data package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(EMLassemblyline)
```


## The data package

A data package forms a collection of data, metadata, and other digital objects that enable understanding, use, and citation of a dataset. Metadata play an important role in understanding each object and how they relate to each other. `EMLassemblyline` works for all digital object formats and has been used by the Environmental Data Initiative (EDI) to create and manage well over 250 data packages for a nearly equal number of research scientists. Editing and revisioning data packages on short notice is made easy by `EMLassemblyline` but requires some forethought to file organization.

## An organization scheme

The organization scheme we’ve found most useful at EDI is a single directory for each data package containing:

*  __data_objects__ A directory of data and other objects to package (e.g. data tables, analysis scripts, .zip files, etc.).
*  __metadata_templates__ A directory of `EMLassemblyline` metadata templates.
*  __eml__ A directory of EML files created by `EMLassemblyline`.
*  __run_EMLassemblyline.R__ An empty R file for scripting the `EMLassemblyline` workflow.

```{r remove_directory, echo=FALSE}

unlink(
  x = '/Users/Colin/Desktop/data_packages/edi_301',
  recursive = TRUE
)

# Set working directory

knitr::opts_knit$set(root.dir = '/Users/Colin/Desktop/data_packages')

```

```{r template_directories, echo=TRUE, paged.print=TRUE, message = FALSE}

# Set working directory

setwd('/Users/Colin/Desktop/data_packages')

# Create the standard data package directory

template_directories(
  path = '.',
  dir.name = 'edi_301'
)

# View directory contents

dir('./edi_301')

```

Any organization scheme is accepted as long as all the template files of a data package are located in the same directory.